UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-06-03 ms/main.tex
fix some text, add more FEV1 stuff for East Asia

Forced expiratory volume in 1 second (FEV1) (figures~\ref{fig:supp_ukbb_fev_f}, \ref{fig:supp_ukbb_fev_m}) also shows strong correlations with certain populations --- South Asian, African, and Caribbean --- having considerably lower measurements on average (see figures~\ref{fig:supp_box_fev_f} and \ref{fig:supp_box_fev_m} for boxplots and p-values). Notably, there appears to be a juncture in the admixture population, highlighted in figure~\ref{fig:supp_comparison_fev_afr}, where the distribution of FEV1 changes. This roughly corresponds to the transition from Black African/Caribbean individuals to those who identified having mixed backgrounds. Boxplots and statistical testing suggest that relative to White British populations, FEV1 values are significantly lower for Black African and Black Caribbean populations, but not for White and Black Caribbean and White and Black African populations. Figure~\ref{fig:supp_comparison_fev_chi_eur} also shows an important difference in FEV1 between self-identified Chinese and a nearby cluster enriched in individuals born in Japan. \sgcomment{boxblot}.

@@

>>>>

Forced expiratory volume in 1 second (FEV1) (figures~\ref{fig:supp_ukbb_fev_f}, \ref{fig:supp_ukbb_fev_m}) also shows strong correlations with certain populations --- South Asian, African, and Caribbean --- having considerably lower measurements on average (see figures~\ref{fig:supp_box_fev_f} and \ref{fig:supp_box_fev_m} for boxplots and p-values). Notably, there appears to be a juncture in the admixture population, highlighted in figure~\ref{fig:supp_comparison_fev_afr}, where the distribution of FEV1 changes. This roughly corresponds to the transition from Black African/Caribbean individuals to those who identified having mixed backgrounds. Boxplots and statistical testing suggest that relative to White British populations, FEV1 values are significantly lower for Black African and Black Caribbean populations, but not for White and Black Caribbean and White and Black African populations. Figure~\ref{fig:supp_comparison_fev_chi_eur} also shows an important difference in FEV1 between self-identified Chinese and a nearby cluster enriched in individuals born in Japan\ref{fig:supp_fev_ridgeplots}.

@@

\begin{figure*}
    \centering
    \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{images/east_asia_ridgeplots_fev.png}
    \end{subfigure}
    \caption{Plots of the distributions of FEV1 by sex for East Asian populations. Individuals were limited to those in the "Chinese/Other Ethnic Group" cluster from figure~\ref{fig:supp_comparison_fev_chi_eur}. The Chinese individuals were those who self-identified their ethnic background as Chinese, and the remaining populations were determined based on country of birth; the categorizations are mutually exclusive. Asterisks indicate significant difference from the Japanese population, using Welch's unpaired t-test with a Bonferroni correction for 3 groups. The dashed lines are the means of the distributions and Japanese populations have consistently higher means, though the test between Chinese and Japanese males fell short of the significance threshold.}
    \label{fig:supp_fev_ridgeplots}
\end{figure*}
